Uncertainty-aware graph-based multimodal remote sensing classification

### 2021-07-06 latex/CEUR/paper.tex
Removed section explaining dropout inference, and will have to assume reader will follow references if unfamiliar.

In order assess the relative ability of evidential uncertainties to express different types of doubt, three commonly used probabilistic measures of uncertainty (aleatoric, epistemic and entropy) are computed to provide comparisons.
For a Bayesian model with inputs, outputs and parameters denoted $x$, $y$ and $\boldsymbol{\theta}$ respectively, the posterior which will be approximated is $p(\boldsymbol{\theta}\vert\mathcal{D})$, where $\mathcal{D}$ is the observed dataset.
The total entropy, made up of both aleatoric and epistemic uncertainty \cite{Depeweg2018} is given by
\begin{equation}
    \mathcal{H}\left(\mathbb{E}_{p(\boldsymbol{\theta}\vert \mathcal{D})}\left[p(y\vert x;\boldsymbol{\theta}\right]\right),
\end{equation}
where $\mathcal{H}()$ computes the differential entropy of a distribution.
Aleatoric uncertainty, which accounts for noise inherent to the data observations \cite{Kendall2017}, is given by
\begin{equation}
    \mathbb{E}_{p\left(\boldsymbol{\theta}\vert \mathcal{D}\right)}\left[\mathcal{H}(p(y\vert x; \boldsymbol{\theta}))\right].
\end{equation}
The epistemic uncertainty, therefore, is given by the difference between these two quantities
\begin{equation}
    \mathcal{H}\left(\mathbb{E}_{p(\boldsymbol{\theta}\vert \mathcal{D})}\left[p(y\vert x;\boldsymbol{\theta}\right]\right)-\mathbb{E}_{p\left(\boldsymbol{\theta}\vert \mathcal{D}\right)}\left[\mathcal{H}(p(y\vert x; \boldsymbol{\theta}))\right].
\label{eq::epistemic}
\end{equation}
In order to compute the above expressions for aleatoric and epistemic uncertainty, the posterior distribution over models is approximated by samples obtained using dropout inference \cite{Gal2016} as is described in Section \ref{sec::dropout_bayesian}.

@@

\subsection{Subjective models}
In order to obtain models which will give node-level Dirichlet distribution parameters, the softmax output activation function used in a GCN is substituted for a ReLU function.
In this way, the model is trained to output non-negative evidence contributions, $\mathbf{E}\in\mathbb{R}^{N\times K}$, where $\mathbf{E}_{i} = \boldsymbol{\alpha}_{i}-\mathbf{1}$ and $\boldsymbol{\alpha}_{i}$ refers to the K-dimensional concentration parameters of the $\text{i}^\text{th}$ node.
In order to train an evidential model, the loss function is made up of two components; a squared error term which is minimised in order to classify a greater proportion of the nodes correctly and a variance term which is minimized to incentivize the model to provide confident predictions where possible.
This loss, $\mathcal{L}(\boldsymbol{\theta})$, is given by
\begin{equation}
\begin{split}
    \mathcal{L}(\boldsymbol{\theta})&=\sum_{i\in\mathbb{L}}\sum_k \left[(p_{ik}-y_{ik})^2+\text{Var}(p_{ik})\right],\\
    &=\sum_{i\in\mathbb{L}}\sum_k \left[(p_{ik}-y_{ik})^2+\frac{\left(\frac{\alpha_{ik}}{S_i}\right)\left(1-\frac{\alpha_{ik}}{S_i}\right)}{S_i-K}\right],
\end{split}
\label{eq:core_loss}
\end{equation}
where $i\in\mathbb{L}$ refers to the fact that the loss is computed using a sum only over nodes in the training set, $\mathbb{L}$.
Models trained with such an output activation and loss function will be denoted using the `S-' prefix in order to indicate they provide subjective predictions, e.g. S-GCN.
\subsection{Dropout inference as a Bayesian approximation}
\label{sec::dropout_bayesian}
It has been shown that a model with dropout can be considered to represent the posterior distribution over models \cite{Gal2016}.
The dropouts are used to represent stochasticity in the model parameters \cite{Ryu2019}.
A given combination of dropouts can be considered to represent a model sampled from the approximate posterior, $q(\boldsymbol{\theta})$, with the actual posterior distribution of models being $p(\boldsymbol{\theta}|\mathcal{D})$.
Monte Carlo sampling can be used to approximately compute quantities such as the term $\mathbb{E}_{p\left(\boldsymbol{\theta}\vert \mathcal{D}\right)}\left[\mathcal{H}(p(y\vert x; \boldsymbol{\theta}))\right]$ in (\ref{eq::epistemic}) allowing the aleatoric and epistemic uncertainty to be calculated.
This can be done using the approximation 
\begin{equation}
    \mathbb{E}_{p\left(\boldsymbol{\theta}\vert \mathcal{D}\right)}\left[\mathcal{H}(p(y\vert x; \boldsymbol{\theta}))\right]\approx\frac{1}{M}\sum_{m=1}^M \mathcal{H}\left(p(y\vert x; \boldsymbol{\theta}^m)\right),
\end{equation}
where $\boldsymbol{\theta}^m\sim q(\boldsymbol{\theta})$ is a model sampled by performing dropout inference.
When this technique has been used, the model name represents this using a `B' in its name, e.g. S-BGCN, showing that this Bayesian approximation has been used.
The Drop-GCN model, which is used for benchmarking the subjective methods, refers to the same technique when applied to a `vanilla' GCN.

>>>>

In order assess the relative ability of evidential uncertainties to express different types of doubt, three commonly used probabilistic measures of uncertainty (aleatoric, epistemic and entropy) are computed to provide comparisons.
For a Bayesian model with inputs, outputs and parameters denoted $x$, $y$ and $\boldsymbol{\theta}$ respectively, the posterior which will be approximated is $p(\boldsymbol{\theta}\vert\mathcal{D})$, where $\mathcal{D}$ is the observed dataset.
The total entropy, made up of both aleatoric and epistemic uncertainty \cite{Depeweg2018} is given by
\begin{equation}
    \mathcal{H}\left(\mathbb{E}_{p(\boldsymbol{\theta}\vert \mathcal{D})}\left[p(y\vert x;\boldsymbol{\theta}\right]\right),
\end{equation}
where $\mathcal{H}()$ computes the differential entropy of a distribution.
Aleatoric uncertainty, which accounts for noise inherent to the data observations \cite{Kendall2017}, is given by
\begin{equation}
    \mathbb{E}_{p\left(\boldsymbol{\theta}\vert \mathcal{D}\right)}\left[\mathcal{H}(p(y\vert x; \boldsymbol{\theta}))\right].
\end{equation}
The epistemic uncertainty, therefore, is given by the difference between these two quantities
\begin{equation}
    \mathcal{H}\left(\mathbb{E}_{p(\boldsymbol{\theta}\vert \mathcal{D})}\left[p(y\vert x;\boldsymbol{\theta}\right]\right)-\mathbb{E}_{p\left(\boldsymbol{\theta}\vert \mathcal{D}\right)}\left[\mathcal{H}(p(y\vert x; \boldsymbol{\theta}))\right].
\label{eq::epistemic}
\end{equation}
In order to compute the above expressions for aleatoric and epistemic uncertainty, the posterior distribution over models is approximated by samples obtained using dropout inference \cite{Gal2016}.

@@

\subsection{Subjective models}
In order to obtain models which will give node-level Dirichlet distribution parameters, the softmax output activation function used in a GCN is substituted for a ReLU function.
In this way, the model is trained to output non-negative evidence contributions, $\mathbf{E}\in\mathbb{R}^{N\times K}$, where $\mathbf{E}_{i} = \boldsymbol{\alpha}_{i}-\mathbf{1}$ and $\boldsymbol{\alpha}_{i}$ refers to the K-dimensional concentration parameters of the $\text{i}^\text{th}$ node.
In order to train an evidential model, the loss function is made up of two components; a squared error term which is minimised in order to classify a greater proportion of the nodes correctly and a variance term which is minimized to incentivize the model to provide confident predictions where possible.
This loss, $\mathcal{L}(\boldsymbol{\theta})$, is given by
\begin{equation}
\begin{split}
    \mathcal{L}(\boldsymbol{\theta})&=\sum_{i\in\mathbb{L}}\sum_k \left[(p_{ik}-y_{ik})^2+\text{Var}(p_{ik})\right],\\
    &=\sum_{i\in\mathbb{L}}\sum_k \left[(p_{ik}-y_{ik})^2+\frac{\left(\frac{\alpha_{ik}}{S_i}\right)\left(1-\frac{\alpha_{ik}}{S_i}\right)}{S_i-K}\right],
\end{split}
\label{eq:core_loss}
\end{equation}
where $i\in\mathbb{L}$ refers to the fact that the loss is computed using a sum only over nodes in the training set, $\mathbb{L}$.
Models trained with such an output activation and loss function will be denoted using the `S-' prefix in order to indicate they provide subjective predictions, e.g. S-GCN.
